Sulfur Dioxide on Water's Surface

### 2011-04-04 Background.tex
Lots of new angle images. Only showing one surface for angle plots instead of both slab surfaces. rewrote the first draft incorporating Steph's comments. Still need an image for the conclusions.

	In this study we create orientational depth-profiles of \suldiox~and \wat. The reference point used for depth calculations is the location of the water surface. A liquid surface, however, is a dynamic boundary that is neither flat nor stationary. To define a water surface location, several previous studies have used the technique of fitting a lineshape to the averaged density profile of the water, and extracting interfacial shape and location parameters.\cite{Shamay2010,Wick2006c,Chowdhary2006} Hyperbolic tangent functions have been used often, and values for the ``Gibb's dividing surface'' location, and interfacial width have thus been determined.\cite{Matsumoto1988} However, in long simulations the location and shape of the interface changes, and the motion of surface waters alters the interfacial width at any given timestep. Thus, the density profile fitting will capture averaged widths and locations, not instantaneous values. Similarly, the averaged values of location and width will obscure information about any drift or deformations the surface undergoes. In this work we approach the problem of defining surface location by calculating a reference location at each timestep by averaging the positions of the waters contained in the topmost monolayer. This provides us with a consistent and intuitive reference point in the simulations to which we relate our analyses, but does not increase the computational burden.

@@

Figure \ref{fig:density-flaw} demonstrates the problem of surface location drift during a simulation. The figure shows a density profile of water and \suldiox~over the course of one of the 10 ns trajectories used in this work. The water density profile (dashed line) is produced by averaging the instantaneous density profile at each timestep in the simulation. The water profile was then fit to a $\tanh$ function (black) to extract the position and width parameters of the water surface. The fitted Gibbs dividing surface has a ``90-10'' width of 8.28 $\AA$. A bulge in the gas-phase (left) side of the water profile is indicative of the drift of the water slab over the course of the trajectory. Thus the calculated location and width from the $\tanh$ line fit are not accurate over long trajectories for defining a stationary reference point. All depth locations in our analyses are calculated relative to the surface location of each timestep of the trajectories.

\begin{figure}[h!]
	\begin{center}
		\includegraphics[scale=1.0]{images/density/Density+SurfaceLocation-flawed.png}
		\caption{Density profiles of \wat~(dashed) and\suldiox~(red) from a 10 ns simulation of the neat-water system with a single sulfur dioxide. The \wat~density profile was fit using a hyperbolic tangent function to extract location and width data. The location of the outter \wat~monolayer was calculated, and the corresponding location distribution (blue) is also plotted. Long simulations of liquid slabs result in drift of the slab location, and the consequent broadening of the water density profile if the surface location is not determined repeatedly throughout the simulation.}
		\label{fig:density-flaw}
	\end{center}
\end{figure}

The new procedure described above for finding the outter monolayer location was implemented, and the distribution of that location is also plotted in figure \ref{fig:density-flaw} (blue). It is apparent from the surface location distribution that the water slab drifts over the 10 ns, spanning approx. 12 $\AA$. However, the maximum standard deviation of the positions of the waters comprising the surface layer at each timestep is only 1.85 $\AA$.

@@

	A second angle, $\phi$, defines the molecular ``twist'' for certain configurations of the molecule. $\phi$ is the angle formed between the vector normal to the plane of the molecule (molecular y-axis) and the system reference axis. The values of $\phi$ fall in the interval $[-\pi,\pi]$. Given certain values of $\theta$, $\phi$ provides additional information about whether the molecular orientation is ``flat'' to the surface (e.g. the plane of the molecule is aligned with the plane of the surface), or if it is perpendicular. %However, the symmetry of the \wat~and \suldiox~molecules create equivalence between the two H, or O, atoms (in the \wat~or\suldiox, respectively). $\phi=-\pi$ is equivalent to $\phi=\pi$, and both situations are equivalent to $\phi=0$. Because $\phi=-\phi$, the value is reported within the range of $[0,\frac \pi 2]$. 
	However, with $\cos(\theta)$ close to 1 or -1 the symmetry of the plane of the surface results in an isotropic distribution in $\cos(\phi)$. Figure \ref{fig:water-angles} shows the angle definitions relating the body-fixed axes to a system-fixed reference axis.

@@

\begin{figure}[h!]
	\begin{center}
		\includegraphics[scale=1.0]{images/angle-cartoons/wateranglessmall.png}
		\caption{The definition of the angles $\theta$ and $\phi$ used to define molecular orientation of \suldiox~and \wat~relative to a reference axis. $\theta$ is the value of the ``tilt'' of the molecular bisector from the reference axis. $\phi$ is the ``twist'' angle that indicates how flat the molecule is relative to the interfacial plane, and is formed between the vector normal to the molecular plane (the body-fixed y-axis) and the reference axis.}
		\label{fig:water-angles}
	\end{center}
\end{figure}

>>>>

	In this study we create orientational depth-profiles of \suldiox~and \wat. The reference point used for depth calculations is the location of the water surface. A liquid surface, however, is a dynamic boundary that is neither flat nor stationary. To define a water surface location, several previous studies have used the technique of fitting a lineshape to the averaged density profile of the water, and extracting interfacial shape and location parameters.\cite{Shamay2010,Wick2006c,Chowdhary2006} Hyperbolic tangent functions have been used often, and values for the ``Gibb's dividing surface'' location, and interfacial width have thus been determined.\cite{Matsumoto1988} However, in long simulations the location and shape of the interface changes, and the motion of surface waters alters the interfacial width at any given timestep. Thus, the density profile fitting will capture averaged widths and locations, not instantaneous values. Similarly, the averaged values of location and width will obscure information about any drift or deformations the surface undergoes. In this work we define the surface location by calculating a reference location at each timestep by averaging the positions of the waters contained in the topmost monolayer. This provides us with a consistent and intuitive reference point in the simulations to which we relate our analyses, but does not increase the computational burden.

@@

Figure \ref{fig:density-flaw} demonstrates the problem of surface location drift during a simulation. The figure shows a density profile of water and \suldiox~over the course of one of the 10 ns trajectories used in this work. The water density profile (dashed line) is produced by averaging the instantaneous density profile at each timestep in the simulation. The water profile was then fit to a $\tanh$ function (black) to extract the position and width parameters of the water surface. The fitted Gibbs dividing surface has a ``90-10'' width of 8.28 $\AA$. A bulge in the gas-phase (left) side of the water profile is indicative of the drift of the water slab over the course of the trajectory. Thus the calculated location and width from the $\tanh$ line fit are not accurate over long trajectories for defining a stationary reference point. All depth locations in our analyses are calculated relative to the surface location at each timestep of the trajectories.

\begin{figure}[h!]
	\begin{center}
		\includegraphics[scale=1.0]{images/density/Density+SurfaceLocation-flawed.png}
		\caption{Density profiles of \wat~(dashed) and \suldiox~(red) from a 10 ns simulation of the neat-water system with a single sulfur dioxide. The \wat~density profile was fit using a hyperbolic tangent function to extract location and width data. The location of the outer \wat~monolayer was calculated, and the corresponding location distribution (blue) is also plotted. Long simulations of liquid slabs result in drift of the slab location, and the consequent broadening of the water density profile if the surface location is not determined repeatedly throughout the simulation.}
		\label{fig:density-flaw}
	\end{center}
\end{figure}

The new procedure described above for finding the outer monolayer location was implemented, and the distribution of that location is also plotted in figure \ref{fig:density-flaw} (blue). It is apparent from the surface location distribution that the water slab drifts over the 10 ns, spanning approx. 12 $\AA$. However, the maximum standard deviation of the positions of the waters comprising the surface layer at each timestep is only 1.85 $\AA$.

@@

	A second angle, $\phi$, defines the molecular ``twist'' for certain configurations of the molecule. $\phi$ is the angle formed between the vector normal to the plane of the molecule (molecular y-axis) and the system reference axis. The values of $\phi$ fall in the interval $[-\pi,\pi]$. For certain values of $\theta$, $\phi$ provides additional information about whether the molecular orientation is ``flat'' to the surface (e.g. the plane of the molecule is aligned with the plane of the surface), or if it is perpendicular. %However, the symmetry of the \wat~and \suldiox~molecules create equivalence between the two H, or O, atoms (in the \wat~or\suldiox, respectively). $\phi=-\pi$ is equivalent to $\phi=\pi$, and both situations are equivalent to $\phi=0$. Because $\phi=-\phi$, the value is reported within the range of $[0,\frac \pi 2]$. 
	However, with $\cos(\theta)$ close to 1 or -1 the symmetry of the plane of the surface results in an isotropic distribution in $\cos(\phi)$. Figure \ref{fig:water-angles} shows the angle definitions relating the body-fixed axes to a system-fixed reference axis.

@@

\begin{figure}[h!]
	\begin{center}
		\includegraphics[scale=1.0]{images/angle-cartoons/molecular-angles.png}
		\caption{(A) The definition of the angles $\theta$ and $\phi$ used to define molecular orientation of \suldiox~and \wat~relative to a reference axis. (B) $\theta$ is the value of the ``tilt'' of the molecular bisector from the reference axis, with values ranging from -1 to +1 indicating different bisector orientations. (C) $\phi$ is the ``twist'' angle formed between the vector normal to the molecular plane (the body-fixed y-axis) and the reference axis. For certain values of $\theta$, $\phi$ shows how flat the molecule is relative to the interfacial plane.}
		\label{fig:water-angles}
	\end{center}
\end{figure}

### 2011-04-04 DensityResults.tex
Lots of new angle images. Only showing one surface for angle plots instead of both slab surfaces. rewrote the first draft incorporating Steph's comments. Still need an image for the conclusions.

One measure of surface activity is the spatial distribution of molecules around the interfacial region. The density distribution of both \wat~and \suldiox~were calculated for the equilibrium MD simulations, and the results are presented in figure \ref{fig:density} for both the neat-water and saturated simulations. The \suldiox~in the neat-water system remained at the top slab surface. While the \suldiox~equilibrated in the saturated system, both in the bulk and gas phases accumulated at both the top and bottom surfaces.

The density distributions of the \suldiox~were fit to gaussian lineshapes as shown in equation \ref{eq:gaussian} using the common definitions of the variables $A$, $\mu$, and $\sigma$, and given as a function of the distance to the top surface, $z$.

\begin{equation}
  \rho(z)=\frac{A}{\sqrt{2 \pi \sigma^2}} e^{-\frac{(z-\mu)^2}{2\sigma^2}}
  \label{eq:gaussian}
\end{equation}

In the neat-water \suldiox~distribution, $\mu=0.781$, $A=0.109$, and $\sigma=1.684$\angs. Thus, the \suldiox~has a surface affinity when in a low-concentration water environment, and it does not fall further into the bulk or move out of the surface into the gas phase. This surface active behavior was found experimentally by a previous SFG study by our group,\cite{Tarbuck2005,Tarbuck2006} and also verified by the computational simulations and spectral calculations of Dang et al.\cite{Baer2010} It was concluded that a surface \suldiox~hydrate complex forms and modifies the structure of water above the Gibbs dividing surface (GDS) shifting \wat~free-OH oscillators in towards the bulk, creating a stronger hydrogen-bonding network. In a later section of this article we look at the effect of \suldiox~on surface water orientation.

Both saturated \suldiox~surfaces exhibit accumulations of \suldiox. However, unlike the neat-water slab, the saturated slab has a non-zero bulk concentration of \suldiox~that does not adsorb to the surface. The coefficient values for the top surface from eq \ref{eq:gaussian} are $\mu_{top}=2.695$, $A_{top}=7.001$, and $\sigma_{top}=3.203$. The bottom surface values are $\mu_{bottom}=-33.846$, $A_{bottom}=6.618$, and $\sigma_{bottom}=3.151$\angs. The added concentration of \suldiox~creates a layer of molecules bound to the top of the water surface. The center of this distribution is further into the gas phase than for the single-\suldiox~molecule surface. The accumulation of \suldiox~in this classical simulation coincides with the experimental conclusions of \suldiox~surface hydrate complex formation. However, without simulating breakble bonds the chemistry that may pull \suldiox~into the bulk water phase, and the subsequent reaction to form ions is absent. We have confidence in the surface complexation that is exhibited, the effect of the \suldiox~on the surface water molecules, and the accumulation of \suldiox at the interface. However, we do not yet conclude that the specific locations of \suldiox~on top of the water surface are as accurate as those found in the DFT ab initio calculations of Dang et al.\cite{Baer2010} As was pointed out in their computations, the classical potential does not reproduce the bonding of the first hydration shell around \suldiox, and so conclusions about specific complex formation or geometries are best made from the ab initio simulation results reported.

>>>>

One measure of surface activity is the spatial distribution of molecules in the interfacial region. The density distribution of both \wat~and \suldiox~were calculated for the equilibrium MD simulations, and the results are presented in figure \ref{fig:density} for both the neat-water and saturated simulations. The \suldiox~in the neat-water system remained at the top slab surface. The \suldiox~in the saturated system accumulated mostly at the surface, but some residual \suldiox~remained in the bulk.

%The density distributions of the \suldiox~were fit to gaussian lineshapes as shown in equation \ref{eq:gaussian} using the common definitions of the variables $A$, $\mu$, and $\sigma$, and given as a function of the distance to the top surface, $z$.

%\begin{equation}
%  \rho(z)=\frac{A}{\sqrt{2 \pi \sigma^2}} e^{-\frac{(z-\mu)^2}{2\sigma^2}}
%  \label{eq:gaussian}
%\end{equation}

In the neat-water slab, the \suldiox~distribution shows a surface affinity when in a low-concentration environment. It does not reside in the bulk or move out of the surface into the gas phase. This surface active behavior was found experimentally by the SFG study by our group,\cite{Tarbuck2005,Tarbuck2006} and also verified by the computational simulations and spectral calculations of Baer et al.\cite{Baer2010} The water behavior hypothesized from the experimental analysis is that a surface \suldiox~hydrate complex forms, and it modifies the structure of water above the Gibbs dividing surface (GDS). \wat~free-OH oscillators are thus shifted in towards the bulk, creating a stronger hydrogen-bonding network. In a later section of this article we look at the effect of \suldiox~on surface water orientation.

Both saturated \suldiox~surfaces exhibit accumulations of \suldiox. However, unlike the neat-water slab, the saturated slab has a non-zero bulk concentration of \suldiox~that does not adsorb to the surface. The added concentration of \suldiox~creates a layer of molecules bound to the top of the water surface. The center of the \suldiox~density distribution is further into the gas phase than for the neat-water surface with a single-\suldiox~molecule. The accumulation of \suldiox~in this classical simulation coincides with the experimental conclusions of \suldiox~surface hydrate complex formation. However, without simulating breakble bonds the chemistry that may pull \suldiox~into the bulk water phase, and the subsequent reaction to form ions is absent. We have confidence in the surface complexation that is exhibited, the effect of the \suldiox~on the surface water molecules, and the accumulation of \suldiox~at the interface. However, we do not yet conclude that the specific locations of \suldiox~on top of the water surface are as accurate as those found in the DFT ab initio calculations of Dang et al.\cite{Baer2010} As was pointed out in their computations, the classical potential does not reproduce the bonding of the first hydration shell around \suldiox, and so conclusions about specific complex formation or geometries are best made from the ab initio simulation results reported.

### 2011-04-04 EquilibratedOrientation.tex
Lots of new angle images. Only showing one surface for angle plots instead of both slab surfaces. rewrote the first draft incorporating Steph's comments. Still need an image for the conclusions.

Molecular orientation analyses were performed to characterize the net molecular orientation of \wat~and \suldiox~molecules in the different locations of the water surface. Histograms were created to show depth-profiles of the angles $\theta$ and $\phi$ in the neat-water and saturated equilibrium systems.

@@

The orientation profiles for \wat~are shown in figure \ref{fig:water-orientation} for both the neat-water (top) and saturated (bottom) systems during the equilibrium MD simulations. In both systems an orientational preference is found at the slab surfaces where the water is furthest towards the gas phase. The histograms are arranged with the plots of $\cos(\theta)$ on the left and $\cos(\phi)$ on the right columns. The top surface is located at 0\angs, and the bottom surface is approx. 30\angs below. It is also noted that $\theta$ values of the bottom surface are inverted with respect to the top surface, mirroring those results.

%The bisector tilt concentrates around $\cos(\theta)=0$ within the first few\angs of the surface, and then the distribution becomes isotropic further into the water bulk. As the tilt nears $\cos(\theta)=0$ the \wat~bisector lies within the plane of the surface indicating a water orientation either flat on the surface, or with some amount of ``twist'' sending the OH bonds in towards, or out of the bulk. The value of $\phi$ determines the ``twist'' in this case. Both systems show a peak in the distributions around $\cos(\phi)=1$ at the water surfaces. This results from an orientation of the water's y-axis (normal to the molecular plane) aligned perpendicular to the plane of the water surface.

The histograms show overall similarities for both systems in their shapes and intensities. The presence of a saturated layer of adsorbed \suldiox~molecules alters the water orientation, but not necessarily the orientational depth of the interface. In both $\theta$ distributions, the orientations of waters in the bulk region are isotropic until 5\angs~below the surface location. At 5\angs~below the surface the water molecules begin to orient with their bisectors within the plane of the interface (perpendicular to the surface normal, $\cos(\theta)\approx 0$). The corresponding location in the plot of $\phi$ shows that those molecules are also mostly flat to the surface ($\cos(\phi)\approx 1$). Moving further out from the bulk and into the gas phase, the distributions show $\cos(\theta)$ increasing. Waters further out from the bulk have fewer bonding interactions, and orient with their hydrogens more towards the gas phase. The bisector pointing further into the gas phase leads to isotropy in the values of $\phi$.

The change in $\cos(\theta)$ is not as rapid at the neat-water surface as in the saturated system. Furthermore, the noise in the distribution above 5\angs~(and below approx -34\angs) in the neat-water system is a result of fewer waters venturing beyond those extents and thus less data far from the surface. This is one indication that waters near a layer of adsorbed \suldiox~venture further above the interface and interact with the adsorbed gas molecules. The interactions with neighboring \suldiox~molecules allow the waters above the surface to orient more perpendicular to the interface. Our experimental SFG studies also predicted this reorienting behavior due to the \suldiox~interactions with the topmost surface waters.\cite{Ota2011}

The distribution of $\cos(\phi)$ is more sharply defined (i.e. less isotropic) for the neat-water system than for the saturated one. Waters on the neat surface lie flatter, whereas the presence of the \suldiox~allows a greater range of ``twist'' for those waters in the plane of the interface. The $\phi$ distributions quickly become isotropic above a couple angstroms from the surface as the bisectors orient more perpendicularly, and below the surface as the bulk water loses any orientational preference.

@@

\begin{figure}[h!]
	\begin{center}
		\includegraphics[scale=1.0]{images/h2o-angles/h2oangles.png}
		\caption{Molecular orientation histograms of \wat~throughout the surface equilibrated systems. The top surface is located at a distance of 0 with negative positions in the bulk of the slab. The bottom slab surface is approx. 30\angs below the top surface. Shown are the angle distributions for $\theta$ (left column) and $\phi$ (right column) in both the neat-\wat~system (top row) and the saturated system (bottom row). The distributions are normalized to account for the changing number of water molecules at different positions in the system.}
		\label{fig:water-orientation}
	\end{center}
\end{figure}

@@

The angular distribution of the single \suldiox~is concentrated primarily in $\cos(\theta)>0$, such that the bisector points out of the water surface. This same distribution occurs in the saturated system for positions below 0\angs, but the orientation becomes isotropic for molecules located above 0\angs, on top of the water surface. Promixity to the water highly orients the \suldiox~with the sulfur atom pointing in towards the water bulk.

>>>>

Molecular orientation analyses were performed to characterize the net molecular orientation of \wat~and \suldiox~molecules in the different locations of the water surface. At each distance from the surface location, an orientation profile was created for both the \wat~and \suldiox~molecules. The orientation distribution for the angles $\theta$ and $\phi$ at all depths were combined to form 2D intensity plots that show how the molecular orientation distributions changed with distance to the surface location. These plots allow for a visual interpretation of how the net orientations are affected when moving from the gas phase to the surface, and then further into the aqueous bulk.

@@

The orientation depth-profiles for \wat~are shown in figure \ref{fig:water-orientation} for both the neat-water (top) and saturated (bottom) systems during the equilibrium MD simulations. In both systems the strongest orientational preference is found at the slab surfaces where the water is furthest towards the gas phase. The histograms are arranged with the plots of $\cos(\theta)$ on the left and $\cos(\phi)$ on the right columns. The surface is located at 0\angs. The angle distributions from the top and bottom slab surfaces were averaged for all the orientation analyses.

The bisector tilt, $\cos(\theta)$, concentrates around $\cos(\theta)=0$ within the first few\angs~of the surface, and then the distribution becomes isotropic further into the water bulk. As the tilt nears $\cos(\theta)=0$ the \wat~bisector lies within the plane of the surface indicating a water orientation either flat on the surface, or with some amount of ``twist'' sending the OH bonds in towards, or out of the bulk. The value of $\phi$ determines the ``twist'' in this case. Both systems show a peak in the distributions around $\cos(\phi)=1$ at the water surfaces. This results from an orientation of the water's y-axis (normal to the molecular plane) aligned perpendicular to the plane of the water surface. Thus, both the neat-water and saturated systems have waters lying mostly flat to the plane of the interface at a distance of 0\angs.

Although the plots show overall similarities for both the neat-water and saturated systems, the presence of a layer of adsorbed \suldiox~molecules alters the water orientation of the waters furthest from the aqueous bulk. The waters located above 0\angs~are further into the gas phase, and interact with the layer of adsorbed \suldiox~molecules. The \suldiox~layer does not provide the same amount of strong hydrogen bonding and solvation that the waters experience further in to the aqueous phase. The resulting orientation of waters above 0\angs, shown in the saturated plot of $\cos(\theta)$ of figure \ref{h2o-angles}, is with a bisector pointing further into the adsorbed \suldiox~gas layer, and both hydrogens pointing outward from the aqueous bulk. The effect is more pronounced further from the water phase, and above 5\angs~the $\cos(\theta)$ distribution is completely centered around $\cos(\theta)=+1$ (see figure \ref{water-angles}).

%The histograms show overall similarities for both systems in their shapes and intensities from approx. 0\angs~and below. The presence of a saturated layer of adsorbed \suldiox~molecules, however, alters the water orientation, but not necessarily the orientational depth of the interface. In both $\theta$ distributions, the orientations of waters in the bulk region are isotropic until 5\angs~below the surface location. At 5\angs~below the surface the water molecules begin to orient with their bisectors within the plane of the interface (perpendicular to the surface normal, $\cos(\theta)\approx 0$). The corresponding location in the plot of $\phi$ shows that those molecules are also mostly flat to the surface ($\cos(\phi)\approx 1$). Moving further out from the bulk and into the gas phase, the distributions show $\cos(\theta)$ increasing. Waters further out from the bulk have fewer bonding interactions, and orient with their hydrogens more towards the gas phase. The bisector pointing further into the gas phase leads to isotropy in the values of $\phi$.

The change in $\cos(\theta)$ is not as rapid at the neat-water surface as in the saturated system. Furthermore, the noise in the distribution above 5\angs~in the neat-water system is a result of fewer waters venturing beyond those extents and thus less data far from the surface. This is one indication that waters near a layer of adsorbed \suldiox~venture further above the interface and interact with the adsorbed gas molecules. The interactions with neighboring \suldiox~molecules allow the waters above the surface to orient more perpendicular to the interface. Our experimental SFG studies also predicted this reorienting behavior due to the \suldiox~interactions with the topmost surface waters.\cite{Ota2011}

The distribution of $\cos(\phi)$ is more sharply defined (i.e. less isotropic) for the neat-water system than for the saturated one. Waters on the neat surface lie flatter, whereas the presence of the \suldiox~allows a greater range of ``twist'' for those waters in the plane of the interface. The $\phi$ distributions quickly become isotropic above the surface as the bisectors orient more perpendicularly, and below the surface as the bulk water loses any orientational preference.

@@

\begin{figure}[h!]
	\begin{center}
		\includegraphics[scale=1.0]{images/h2o-angles/h2oangles.png}
		\caption{Molecular orientation histograms of \wat~throughout the surface equilibrated systems. The top surface is located at a distance of 0 with negative positions in the bulk of the slab. The bottom slab surface is approx. 30\angs~below the top surface. Shown are the angle distributions for $\theta$ (left column) and $\phi$ (right column) in both the neat-\wat~system (top row) and the saturated system (bottom row). The distributions are normalized to account for the changing number of water molecules at different positions in the system.}
		\label{fig:water-orientation}
	\end{center}
\end{figure}

@@

Within 5\angs~of the water surface (the approximate distance of interaction with surface water molecules) the angular distribution of the single \suldiox~(in the neat-water system) is concentrated primarily in $\cos(\theta)>0$. The peak of the distribution occurs at $\cos(\theta)=1$. The \suldiox~bisector points out of the water surface, with the sulfur atom pointing towards the aquoues bulk, and the two oxygens pointing into the gas phase. This same distribution occurs in the saturated system for positions below 5\angs, but the orientation becomes isotropic for molecules located above 0\angs, on top of the water surface. After 5\angs~above the surface both distributions become isotropic. Promixity to the water highly orients the \suldiox~with the sulfur atom pointing in towards the water bulk. Moving further away from the water surface, and interacting less with \wat~molcules allows for greater orientational freedom as exhibited in the isotropy of the distributions above 5\angs. Because both the neat-water and saturated systems show rather similar distributions, it is possible that the concentration of \suldiox~does not strongly affect the orientation of the \suldiox, unlike the orientations of waters.

The plots of $\cos(\phi)$ are both isotropic, although the neat-water system plot is quite noisy from the small data set. Because the \suldiox~near the surface is oriented perpendicularly to the interface, the $\cos(\phi)$ orientation is expected to be isotropic. Further from the water surface where the bisector orientation becomes isotropic, the $\cos(\phi)$ behavior is expected to behave similarly. For the \suldiox~orientation, the $\phi$ angle does not provide further information regarding the surface behavior.

### 2011-04-04 Introduction.tex
Lots of new angle images. Only showing one surface for angle plots instead of both slab surfaces. rewrote the first draft incorporating Steph's comments. Still need an image for the conclusions.

Interactions between sulfur species and aqueous systems are the current topic of an active area of research. Sulfur dioxide remains an important industrial product, and also enters the environment naturally through terrestrial processes. \suldiox~acts as a major component of atmospheric pollution, and is a precursor to acid rain formation, and cloud nucleation. Its high solubility in water makes \suldiox~ an integral compound in many atmospheric reactions and surface reaction mechanisms. We seek a fuller understanding of the interactions between \suldiox~and water interfaces because of the central role \suldiox~takes in many atmospheric particle and aerosol reactions.

Our recent experimental works provided insight into the adsorption and reaction of gases in atmospheric aerosols.\cite{Ota2011} We concluded that an \suldiox~surface hydrate complex forms when an aqueous surface is exposed to \suldiox~gas. A computational study by Dang et al.\cite{Baer2010} then made a series of predictions of the specific nature of the hydrated complex through classical and ab initio simulations. That work developed a detailed picture of the nature of the \suldiox~surface complex with water, and related it to the surface water OH vibrational IR spectra.

Our latest experimental work on \suldiox~behavior further expands our understanding of \suldiox~in atmospheric processes.\cite{Ota2011} A temperature study showed that the binding of gaseous \suldiox~to water interfaces at atmospherically relevant cold temperatures is greatly enhanced. The increased binding facilitates the reaction with water to produce dissolved sulfur species. However, the surface binding of unreacted \suldiox~complexes was found to be a completely reversible process. The same study also showed that low pH aqueous solutions, as often found in atmospheric aerosol composition, inhibit the reaction of \suldiox, but do not affect the surface binding of the gas molecules. The resultant experimental spectra suggest that the surface waters reorient in response to the presence of gaseous \suldiox, with the interfacial water ``free-OH'' bonds orienting more perpendicularly to the plane of the aqueous surface.

>>>>

Interactions between sulfur species and aqueous systems are the current topic of an active area of research. Sulfur dioxide remains an important industrial product, and also enters the environment naturally through terrestrial processes. \suldiox~acts as a major component of atmospheric pollution, and is a precursor to acid rain formation, and cloud nucleation. Its high solubility in water makes \suldiox~ an integral compound in many atmospheric reactions and surface reaction mechanisms. We seek a fuller understanding of the interactions between \suldiox~and water at interfaces because of the central role \suldiox~takes in many atmospheric particle and aerosol reactions.

Our recent experimental works provided insight into the adsorption and reaction of gases in atmospheric aerosols.\cite{Tarbuck2005,Tarbuck2006} We concluded that an \suldiox~surface hydrate complex forms when an aqueous surface is exposed to \suldiox~gas. A computational study by Dang et al.\cite{Baer2010} then made a series of predictions of the specific nature of the hydrated complex through classical and ab initio simulations. That work developed a detailed picture of the nature of the \suldiox~surface complex with water, and related it to the surface water OH vibrational IR spectra.

Our latest experimental work on \suldiox~behavior further expands our understanding of \suldiox~in atmospheric processes.\cite{Ota2011} A temperature study showed that the binding of gaseous \suldiox~to water surfaces at atmospherically relevant cold temperatures is greatly enhanced. The increased binding facilitates the reaction with water to produce dissolved sulfur species. However, the surface binding of unreacted \suldiox~complexes was found to be a completely reversible process. The same study also showed that low pH aqueous solutions, as often found in atmospheric aerosol composition, inhibit the reaction of \suldiox, but do not affect the surface binding of the gas molecules. The resultant experimental spectra suggest that the surface waters reorient in response to the presence of gaseous \suldiox. From the fitting analysis, one possible geometry of water was posited with the interfacial water ``free-OH'' bonds orienting more perpendicularly to the plane of the aqueous surface.

### 2011-04-04 Methods.tex
Lots of new angle images. Only showing one surface for angle plots instead of both slab surfaces. rewrote the first draft incorporating Steph's comments. Still need an image for the conclusions.

% need to add in the specifics about the SMD routine, force constants, calculations, etc.
A second set of simulations began with an equilibrated water slab as in the surface equilibrated method above. However, in both the neat-water and saturated systems equilibrated systems, a single \suldiox~was introduced 20\angs~above the water slab surface, with the sulfur atom tethered to its initial position. The systems were evolved for 1 ns, taking coordinate snapshots every 20 ps to create 50 starting points for further simulations. Steered molecular dynamics (SMD) were then performed on the 50 system configurations (in both the neat-water and saturated configurations) to guide the \suldiox~down towards a tethered water 15\angs under the water slab surface by applying a small steering force to the \suldiox-sulfur atom.\cite{Isralewitz2001} The \suldiox~thus passed through the continuum of environments from gas phase to (neat- and saturated) water surface adsorption, and finally absorption into the bulk of the \wat~slab. Each of the SMD simulations were performed for a total of 200 ps, using a timestep of 1 fs, and taking snapshots of the system every 25 fs. Figure \ref{fig:starting-configurations} illustrates two sample starting configurations for the SMD simulations, showing both the neat-water slab and the saturated slab configurations before steering the \suldiox~towards the water bulk.

A separate set of SMD simulations were performed with tethering to one of the \suldiox-oxygens. This was done to ensure that the orientation of the \suldiox~during the adsorption transit was not an artifact of the choice of atom used for tethering. The simulations produced the same results (not shown) for the orientational analyses, so the data from the original tethering scheme was used.

>>>>

% need to add in the specifics about the SMD routine, force constants, calculations, etc.
A second set of simulations began with an equilibrated water slab as in the surface equilibrated method above. However, in both the equilibrated neat-water and saturated systems, a single \suldiox~was introduced 20\angs~above the water slab surface, with the sulfur atom tethered to its initial position. The systems were evolved for 1 ns, taking coordinate snapshots every 20 ps to create 50 starting points for further simulations. Steered molecular dynamics (SMD) were then performed on the 50 system configurations (in both the neat-water and saturated configurations) to guide the \suldiox~down towards a tethered water 15\angs~under the water slab surface by applying a small steering force to the \suldiox-sulfur atom.\cite{Isralewitz2001} The \suldiox~thus passed through the continuum of environments from gas phase to (neat- and saturated) water surface adsorption, and finally absorption into the bulk of the \wat~slab. Each of the SMD simulations were performed for a total of 200 ps, using a timestep of 1 fs, and taking snapshots of the system every 25 fs. Figure \ref{fig:starting-configurations} illustrates two sample starting configurations for the SMD simulations, showing both the neat-water slab and the saturated slab configurations before steering the \suldiox~towards the water bulk.

A separate set of SMD simulations were performed with tethering to one of the \suldiox-oxygens. This was done to ensure that the orientation of the \suldiox~during the adsorption transit was not an artifact of the choice of atom used for tethering. The simulations produced the same results (not shown) for the orientational analyses, so the data from the original tethering scheme were used.
